Simulations of Ballot Polling Risk-Limiting Audits

### 2023-02-28 providence/sections/pilot.tex
I'm done

A first round size of $140$ ballots with large probability of stopping ($0.95$) was selected.
Selection order was tracked for the sake of analysis.
As expected, the audit concluded in the first round. The \Providence risk measure was $0.0418$. This is the smallest risk for which the sample would have passed a \Providence RLA for the announced election---the p-value of the statistical test and the inverse of the \Providence ratio defined in section \ref{sec:prov}. Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Providence, \Minerva and \BRAVO (both EoR and SO), all similarly defined.

\begin{table}[h!]
\begin{center}
\begin{tabular}{ |c|c|c|c|c| } 
\hline
%\diagbox[dir=NW]{First \\Round \\Size}{RLA}
ballots& \rotatebox{45}{\Providence} & \rotatebox{45}{\Minerva} & \rotatebox{45}{SO \BRAVO} & \rotatebox{45}{EoR \BRAVO} \\
\hline
140 & \bf{0.0418} & \bf{0.0418} & \bf{0.0541} & 0.366 \\
\hline
\end{tabular}
\end{center}
\caption{Risk measures for the drawn first round of $140$ ballots in the RI pilot audit. Risks in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
\label{tab:pilot-risks}
\end{table}

@@

We observe two interesting phenomena in this particular sample's selection order. 
First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the winner votes in the sample (orange line) surpasses the minimum number of winner ballots need to meet the \BRAVO stopping condition (blue line) earlier in the sample\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.}. EoR \BRAVO, however, does not stop at sample size $140$, the number of relevant ballots drawn during the pilot. It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but the rest of the sample is ignored. Second, the orange line is below the dotted green line, which represents half the sample size; only $5$ of the first $11$ ballots were for the announced winner. A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size. 
%Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{pilot_sequence.png}
\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner is shown. Observe that SO \BRAVO would stop because the minimum number of ballots required to satisfy the \BRAVO stopping condition (blue line) is achieved early on (when the orange line crosses the blue one) in the audit. However, because the \BRAVO condition is no longer satisfied at $140$ ballots, an EoR \BRAVO audit would not have stopped. Further, at a sample size of $11$, the orange line is not even above the dotted green lime representing half the sample size and the winner has fewer than half the relevant votes in the sample.}
\label{fig:pilot_sequence}
\end{figure}

>>>>

A first round size of $140$ ballots with large probability of stopping ($0.95$) was selected.
Selection order was tracked for the sake of analysis.
As expected, the audit concluded in the first round. The \Providence risk measure was $0.0418$. This is the smallest risk for which the sample would have passed a \Providence RLA for the announced election---the p-value of the statistical test and the inverse of the \Providence ratio $\omega$ of Definition \ref{def:minervatwo}. Table~\ref{tab:pilot-risks} shows risk measures for the drawn sample using \Providence, \Minerva and \BRAVO (both EoR and SO), all similarly defined.

\begin{table}[h!]
\begin{center}
\begin{tabular}{ |c|c|c|c|c| } 
\hline
%\diagbox[dir=NW]{First \\Round \\Size}{RLA}
ballots& \rotatebox{45}{\Providence} & \rotatebox{45}{\Minerva} & \rotatebox{45}{SO \BRAVO} & \rotatebox{45}{EoR \BRAVO} \\
\hline
140 & \bf{0.0418} & \bf{0.0418} & \bf{0.0541} & 0.366 \\
\hline
\end{tabular}
\end{center}
\caption{Risk measures for the drawn first round of $140$ ballots in the RI pilot audit. Risk measures in bold meet the risk-limit ($10\%$) and thus correspond to audits that would stop.}
\label{tab:pilot-risks}
\end{table}

@@

We observe two interesting phenomena in this particular sample's selection order. 
First, an SO \BRAVO audit of this sample stops because the \BRAVO condition is met when the winner votes in the sample (orange line) surpasses the minimum number of winner ballots need to meet the \BRAVO stopping condition (blue line) earlier in the sample\footnote{Such cases also provide insight into how \Providence is a sharper test in expectation because SO \BRAVO ignores information from the rest of the sample after the \BRAVO condition is met at some point earlier in the selection order.}. EoR \BRAVO, however, does not stop at sample size $140$, the number of relevant ballots drawn during the pilot. It might be difficult to explain to the public why SO \BRAVO stops in more extreme cases like this, where the condition is met early in the sample, but the rest of the sample is ignored. Second, the orange line is below the dotted green line, which represents half the sample size, at a sample size of $11$; only $5$ of the first $11$ ballots were for the announced winner. A first round of size $11$ would have resulted in a smaller average total number of ballots drawn, but would have provided a misleading sample (suggesting that the winner was incorrectly reported) due to a too-small sample size. 
%Both these ideas are addressed more thoroughly in section~\ref{sec:workload}.

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{pilot_sequence.png}
\caption{For each sample size from $1$ to $140$, the intermediate cumulative sum of ballots for the announced winner is shown. Observe that SO \BRAVO would stop because the minimum number of ballots required to satisfy the \BRAVO stopping condition (blue line) is achieved early on (when the orange line crosses the blue one) in the audit. However, because the \BRAVO condition is no longer satisfied at $140$ ballots, an EoR \BRAVO audit would not have stopped. Further, at a sample size of $11$, the orange line is not even above the dotted green line representing half the sample size and the winner has fewer than half the relevant votes in the sample.}
\label{fig:pilot_sequence}
\end{figure}
